---
title: "Study Design Review"
author: "Omar Elashkar"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Study Design Review}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  fig.width = 15,
  fig.height = 10, 
  # use toc 
  toc = TRUE

)
```

Pharmacokinetic studies require careful planning and design to ensure that the data collected is suitable for analysis. There are few common studies with primary focus on PK such as : 


```{r}
library(PKbioanalysis)
library(rxode2)
```


## Single Ascending Dose (SAD)
SAD used for assessing MTD, clinical PK and dose proportionality. Each dose level is tested in different cohort of subjects.


```{r}
ev <- et(
  id = 1,
  amt = c(100, 200, 300, 400, 500), 
  ii = 24, 
  addl = 4, 
  cmt = "dose", 
  evid = 1, time = 2, arm = 1) |> View()


```

```{r}
sample_times <- seq(0, 120, by = 30) # Sample times in minutes
generate_96() |> 
  add_samples(
    samples = 1:2,
    dose = "",
    sex = "M",
    time = sample_times) |>
  add_samples(
    samples = 3:7,
    dose = "placebo",
    sex = "M",
    time = sample_times) |>
  add_samples(
    samples = 8:9,
    dose = "placebo",
    sex = "M",
    time = sample_times) |>
  add_samples(
    samples = 10:15,
    dose = "placebo",
    sex = "M",
    time = sample_times) |>
    plot()

```


## Multiple Ascending Dose (MAD)
MAD studies are used for steady state PK, accumulation, and long-term safety.
## Bioavailability Study 
## Bioequivalence Study
- Trt A caps vs trt B liquid formulation
### Parallel Design
### Randomized Crossover Study
## Drug-Drug Interaction Study
## Food Effect Study
## Hepatic Impairment Study
## Renal Impairment Study

```{r}
study <- generate_96() |> 
    add_samples(1:5, dose = "A", factor = "M", time = 1:5*30) |>
    add_samples(6:10, dose = "A", factor = "F", time = 1:5*30) |>
    add_samples(11:15, dose = "B", factor = "M", time = 1:5*30) |>
    add_samples(16:18, dose = "B", factor = "F", time = 1:3*30) 

```

```{r}
plot(study)
```



```{r}
data(Indometh)
plate <- generate_96() |> 
  add_samples(
    samples = Indometh$Subject, 
    time = Indometh$time, 
    dose = "100mg",
    route = "IV",
    cmt = "plasma")
```

```{r}
plot(plate, color = "samples")
plot(plate, color = "time")
```

```{r}
plot_design(plate)
```